Clustering in a hyperbolic model of complex networks.

### 2019-02-05 concentration.tex
updated notation and added comments to sections 1-3

\section{Concentration for $c_{\Pcal, n}(k)$ (Proving Proposition \ref{prop:concentration_local_clustering_P_n})}
\label{sec:concentration_c_P_n}

In this section we establish a concentration result for the local clustering function $c_{\Pcal, n}^\ast(k)$ in the finite model $G_{\Pcal,n}$. Similar to the previous section we will focus on those nodes $p = (x,y) \in \Kcal_{C}(k_n)$.

@@

Recall the adjusted triangle count function \eqref{eq:def_tilde_triangle_indicator}
\[
	\widetilde{T}_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal}(p_1) \cap \Rcal_n}.
\]
and the concentration set
\[
	\Kcal_C(k_n) = \left\{p \in \R : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \vee 0 \le e^{\frac{y}{2}}
		\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \wedge e^{R_n/2} \right\},
\]
with
\[
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\]
We will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\Pcal,n}(k, C) = \sum_{p \in \Pcal_n \cap K_C(k)} \ind{D_{\Pcal,n}(p) = k} \widetilde{T}_{\Pcal,n}(p),
\end{equation}
where
\[
	\widetilde{T}_{\Pcal,n}(p) := \sum_{(p_1, p_2) \in \Pcal\setminus (0,y)}^{\ne} \widetilde{T}_{\Pcal,n}(p,p_1,p_2),
\]
where the sum is over all distinct pairs $(p_1, p_2) \in \Pcal \setminus \{(0,y)\}$.

@@

%\begin{lemma}\label{lem:L1_convergence_c_Pcal_n}
%Let $\nu > 0$, $\alpha > \frac{1}{2}$ and $k_n = o\left(n^{\frac{1}{2\alpha + 1}}\right)$. Then, as $n \to \infty$,
%\[
%	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{lemma}
%
%\begin{proof}
%Let $0 < \delta < 1$ and define the following two events
%\begin{align*}
%	A_n &= \left\{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| \le \Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}\right\}\\
%	B_n &= \left\{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| \le n\right\},
%\end{align*}
%so that
%\[
%	1 = \ind{A_n} + \ind{A_n^c}\ind{B_n} + \ind{B_n^c}.
%\]
%Next we note that on the event $A_n$
%\[
%	\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right| 
%	\le \frac{\Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}}{\Exp{N_{\Pcal,n}(k_n)}+\Exp{N_{\Pcal,n}(k_n)}^{\frac{1 + \delta}{2}}}
%	\le \Exp{N_{\Pcal,n}(k_n)}^{-\frac{1 - \delta}{2}},
%\]
%and on the event $B_n$
%\[
%	\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right| 
%	\le \frac{n}{\Exp{N_{\Pcal,n}(k_n)} + n} \le 1.
%\]
%Therefore we have
%\begin{align*}
%	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|}
%	&= \Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|}\\
%	&\le \Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n}}\Exp{N_{\Pcal,n}(k_n)}^{-\frac{1 - \delta}{2}} \\
%	&\hspace{10pt}+ \Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}} \\
%	&\hspace{10pt}+ \Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|\ind{B_n^c}}
%\end{align*}
%Since $\Exp{N_{\Pcal,n}(k_n)} \to \infty$, the first term is clearly $\smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}$. For the third term we have
%\begin{align*}
%	\Exp{c_{\Pcal, n}^\ast(k_n)\left|\frac{\Exp{N_{\Pcal,n}(k_n)}}{N_{\Pcal,n}(k_n)} - 1\right|\ind{B_n^c}} 
%	&= \bigO{\Prob{B_n^c}}
%		= \bigO{\Prob{\left|N_{\Pcal,n}(k_n) - \Exp{N_{\Pcal,n}(k_n)}\right| > n}}\\
%	&= \bigO{e^{-\frac{n^2}{\Exp{N_{k_n}} + n}}} = \bigO{e^{-n}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
%\end{align*}
%Hence we are left to show that
%\begin{equation}\label{eq:L1_convergence_second_term}
%	\Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
%\end{equation}
%By writing
%\begin{align*}
%	c_{\Pcal, n}^\ast(k_n) = \frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%	+ \frac{T_{\Pcal,n}(k_n,\varepsilon)- \widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%\end{align*}
%and using Lemma~\ref{lem:clustering_error_T_term} we get
%\begin{align*}
%	\Exp{c_{\Pcal, n}^\ast(k_n)\ind{A_n^c}\ind{B_n}}
%	&\le \Exp{\frac{(1+\smallO{1})\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%		\ind{A_n^c}} + \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}}.
%\end{align*}
%For the last step we use H\"{o}lder's inequality and Proposition \ref{prop:concentration_tilde_T_P_n} to get
%\begin{align*}
%	\Exp{\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}\ind{A_n^c}}
%	&\le \Exp{\left(\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}
%		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}\right)^2}^{\frac{1}{2}} \Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \left(1 + \smallO{1}\right) 	
%		\frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
%		\Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \bigO{\Exp{c_{\Pcal, n}^\ast(k_n)}}\Prob{A_n^c}^{\frac{1}{2}} = \smallO{\Exp{c_{\Pcal, n}^\ast(k_n)}},
%\end{align*}
%since $\Prob{A_n} = o(1)$. This establishes \eqref{eq:L1_convergence_second_term} and hence we conclude that
%\[
%	\Exp{\left|c_{\Pcal, n}^\ast(k_n) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{proof}

\begin{proof}[Proof of Proposition \ref{prop:concentration_local_clustering_P_n}]
Again, we write
\begin{align*}
	c_{\Pcal, n}^\ast(k_n) = \frac{\widetilde{T}_{\Pcal,n}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	+ \frac{\left(T_{\Pcal,n}(k_n)-\widetilde{T}_{\Pcal,n}(k_n,C)\right)}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}},
\end{align*}
so that by~\eqref{eq:asymptotic_equivalence_contraint_triangle_count},
\begin{equation}\label{eq:concentration_local_clustering_P_n_main}
	\Exp{c_{\Pcal,n}^\ast(k_n)} = \frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	+ \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}}.
\end{equation}
%Therefore, by Lemma \ref{lem:L1_convergence_c_Pcal_n},
%\begin{align*}
%	\Exp{\left|c_{\Pcal,n}(k_n) - \Exp{c_{\Pcal,n}^\ast(k_n)}\right|}
%	&\le \Exp{\left|c_{\Pcal,n}^\ast(k_n) - \Exp{c_{\Pcal,n}^\ast(k_n)}\right|}
%		+ \Exp{\left|c_{\Pcal,n}(k_n) - c_{\Pcal,n}^\ast(k_n)\right|}\\
%	&= \frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
%		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}} + \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}}.
%		\numberthis \label{eq:concentration_local_clustering_P_n_term1}
%\end{align*}
Next, we use Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,C) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)^2} 
		- \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\Pcal,n}(k_n)}}
	&= \smallO{\Exp{c_{\Pcal,n}^\ast(k_n)}},
\end{align*}
which together with~\eqref{eq:concentration_local_clustering_P_n_main} yields the required result.
\end{proof}

\subsection{Joint neighborhoods and degrees in $G_{\Pcal,n}(\alpha,\nu)$}

To prove Proposition~\ref{prop:concentration_tilde_T_P_n} we need to understand the joint degree distribution in $G_{\Pcal,n}$ and subsequently the joint neighborhoods of two points $p, p^\prime \in \Rcal_n$. We perform the analysis in this section. The two main results, which will be the crucial technical ingredients for the proof of Proposition~\ref{prop:concentration_tilde_T_P_n}, are Lemma~\ref{lem:joint_degree_factorization} and Lemma~\ref{lem:joint_degree_distribution_shift}.

@@

We start with analyzing joint neighborhoods in $G_{\Pcal,n}$. Let $p, p^\prime \in \Rcal_n$. Then we denote by $\Ncal_{\Pcal,n}(p\Delta p^\prime)$ the number of disjoint neighbors of $p$ and $p^\prime$ in $G_{\Pcal,n}$, i.e. those points that belong to either $\BallPon{p}$ or $\BallPon{p^\prime}$ but not to their intersection. In addition we denote by $\Ncal_{\Pcal,n}(p,p^\prime)$ the number of joint neighbors of $p$ and $p^\prime$. We shall establish lower bounds on the expect number of these disjoint neighbors $\Exp{\Ncal_{\Pcal,n}(p\Delta p^\prime)}$ as well as an asymptotic expression for the expected number of joint neighbors. For this we will distinguish between the cases where the distance between the $x$-coordinates of $p$ and $p^\prime$ is small or large. Figure~\ref{fig:representation_disjoint_neighborhoods_P_n}-\ref{fig:representation_disjoint_neighborhoods_P_n_3} show the different situations that occur.

@@

	%The box \Rcal_n
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

@@

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $G_{\Pcal,n}(\alpha,\nu)$ when $|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
\label{fig:representation_disjoint_neighborhoods_P_n}
\end{figure}~
\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal_n
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

@@

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $G_{\Pcal,n}(\alpha,\nu)$ when $e^{\frac{y + y^\prime}{2}} < |x - x^\prime| \le e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
\label{fig:representation_disjoint_neighborhoods_P_n_2}
\end{figure}~
\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal_n
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

@@

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $G_{\Pcal,n}(\alpha,\nu)$ when $|x - x^\prime| > e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma \ref{lem:disjoint_neighbors_P_n_large}.}
\label{fig:representation_disjoint_neighborhoods_P_n_3}
\end{figure}

We start by analyzing the shape of the neighborhoods. Due to symmetry and the fact that we have identified the left and right boundaries of the box $\Rcal_n$, we can, without loss of generality, assume that $p = (0,y)$ and $p^\prime = (x^\prime,y^\prime)$ with $x^\prime > 0$ and $y^\prime \le y$. To understand the computation it is helpful to have a picture of the different situations. Figure \ref{fig:representation_disjoint_neighborhoods_P_n} and Figure \ref{fig:representation_disjoint_neighborhoods_P_n_2} show two different situations for small distance in the $x$-coordinates, in which case the number of disjoint neighbors is small. The case where this distance is large, and the number of disjoint neighbors is expected to be large, is show in Figure \ref{fig:representation_disjoint_neighborhoods_P_n_3}. There are several different quantities that are important. The first are the heights $h_1(p^\prime)$ and $h_2(p^\prime)$ where, respectively, the left and right boundaries of the ball $\BallPon{p^\prime}$ go outside the box $\Rcal_n$. Note that when $x = 0$ then these height are the same and we denote this by $h(y)$. We also need to know the coordinates $h(p,p^\prime)$ and $x^\ast(p,p^\prime)$ of the intersection of the right boundary of the neighborhood of $p$ with the left boundary of the neighborhood of $p^\prime$. Finally we will denote by $d(p,p^\prime)$ the distance between the lower right boundary of $\BallPon{p}$ and the lower left of $\BallPon{p^\prime}$, which is positive only when the bottom parts of both neighborhoods do not intersect, compare Figures \ref{fig:representation_disjoint_neighborhoods_P_n} and \ref{fig:representation_disjoint_neighborhoods_P_n_3}. The full expressions of all these functions are given below for further reference.

@@

\begin{lemma}\label{lem:disjoint_neighbors_P_n}
Let $p, p ^\prime \in \Rcal_n$. Then whenever $|x - x^\prime| \le e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$,
\begin{align*}
	\Exp{\Ncal_{\Pcal,n}(p \Delta p^\prime)}
	&\ge \frac{\nu}{\pi}|x^\prime - x|\left(1 - \left(\frac{2}{\pi}\right)^{2\alpha}e^{-\alpha(R_n - y^\ast)}\right)\\
	&\hspace{10pt}+ \xi_{\alpha, \nu}\left|e^{\frac{y}{2}} - e^{\frac{y^\prime}{2}}\right|
		\left(1 - \left(\frac{2}{\pi}\right)^{2\alpha - 1}e^{-(\alpha - \frac{1}{2})(R_n-y^\ast)}\right).
\end{align*}
\end{lemma}

>>>>

\section{Concentration for $c(k; \Gbox)$ (Proving Proposition \ref{prop:concentration_local_clustering_P_n})}
\label{sec:concentration_c_P_n}

In this section we establish a concentration result for the local clustering function $c^\ast(k; \Gbox)$ in the finite model $\Gbox$. Similar to the previous section we will focus on typical points $p = (0,y)$ with $y \in \Kcal_{C}(k_n)$.

@@

Recall the adjusted triangle count function \eqref{eq:def_tilde_triangle_indicator}
\[
	\widetilde{T}_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p}}\ind{p_2 \in \BallPon{p}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal}.
\]
and the concentration set
\[
	\Kcal_C(k_n,n) = \left\{p \in \Rcal : \frac{k_n - C \kappa_n}{\xi} \vee 0 \le e^{\frac{y}{2}}
		\le \frac{k_n + C \kappa_n}{\xi} \right\},
\]
with
\[
	\kappa_n = \sqrt{k_n \log(k_n)}.
\]
We will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\Pcal,n}(k, C) = \sum_{p \in \Pcal_n \cap \Kcal_{C,n}(k_n)} \ind{\text{deg}_{\text{box}}(p) = k} \widetilde{T}_{\Pcal,n}(p),
\end{equation}
where 
\[
	\widetilde{T}_{\Pcal,n}(p) := \sum_{(p_1, p_2) \in \Pcal \setminus p}^{\ne} \widetilde{T}_{\Pcal,n}(p,p_1,p_2),
\]
where the sum is over all distinct pairs $(p_1, p_2) \in \Pcal \setminus p$.

@@

%\begin{lemma}\label{lem:L1_convergence_c_Pcal_n}
%Let $\nu > 0$, $\alpha > \frac{1}{2}$ and $k_n = o\left(n^{\frac{1}{2\alpha + 1}}\right)$. Then, as $n \to \infty$,
%\[
%	\Exp{\left|c^\ast(k_n; \Gbox) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{lemma}
%
%\begin{proof}
%Let $0 < \delta < 1$ and define the following two events
%\begin{align*}
%	A_n &= \left\{\left|N_{\text{box}}(k_n) - \Exp{N_{\text{box}}(k_n)}\right| \le \Exp{N_{\text{box}}(k_n)}^{\frac{1 + \delta}{2}}\right\}\\
%	B_n &= \left\{\left|N_{\text{box}}(k_n) - \Exp{N_{\text{box}}(k_n)}\right| \le n\right\},
%\end{align*}
%so that
%\[
%	1 = \ind{A_n} + \ind{A_n^c}\ind{B_n} + \ind{B_n^c}.
%\]
%Next we note that on the event $A_n$
%\[
%	\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right| 
%	\le \frac{\Exp{N_{\text{box}}(k_n)}^{\frac{1 + \delta}{2}}}{\Exp{N_{\text{box}}(k_n)}+\Exp{N_{\text{box}}(k_n)}^{\frac{1 + \delta}{2}}}
%	\le \Exp{N_{\text{box}}(k_n)}^{-\frac{1 - \delta}{2}},
%\]
%and on the event $B_n$
%\[
%	\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right| 
%	\le \frac{n}{\Exp{N_{\text{box}}(k_n)} + n} \le 1.
%\]
%Therefore we have
%\begin{align*}
%	\Exp{\left|c^\ast(k_n; \Gbox) - c_{\Pcal, n}(k_n)\right|}
%	&= \Exp{c^\ast(k_n; \Gbox)\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right|}\\
%	&\le \Exp{c^\ast(k_n; \Gbox)\ind{A_n}}\Exp{N_{\text{box}}(k_n)}^{-\frac{1 - \delta}{2}} \\
%	&\hspace{10pt}+ \Exp{c^\ast(k_n; \Gbox)\ind{A_n^c}\ind{B_n}} \\
%	&\hspace{10pt}+ \Exp{c^\ast(k_n; \Gbox)\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right|\ind{B_n^c}}
%\end{align*}
%Since $\Exp{N_{\text{box}}(k_n)} \to \infty$, the first term is clearly $\smallO{\Exp{c^\ast(k_n; \Gbox)}}$. For the third term we have
%\begin{align*}
%	\Exp{c^\ast(k_n; \Gbox)\left|\frac{\Exp{N_{\text{box}}(k_n)}}{N_{\text{box}}(k_n)} - 1\right|\ind{B_n^c}} 
%	&= \bigO{\Prob{B_n^c}}
%		= \bigO{\Prob{\left|N_{\text{box}}(k_n) - \Exp{N_{\text{box}}(k_n)}\right| > n}}\\
%	&= \bigO{e^{-\frac{n^2}{\Exp{N_{k_n}} + n}}} = \bigO{e^{-n}} = \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%\end{align*}
%Hence we are left to show that
%\begin{equation}\label{eq:L1_convergence_second_term}
%	\Exp{c^\ast(k_n; \Gbox)\ind{A_n^c}\ind{B_n}} = \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%\end{equation}
%By writing
%\begin{align*}
%	c^\ast(k_n; \Gbox) = \frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%	+ \frac{T_{\Pcal,n}(k_n,\varepsilon)- \widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%\end{align*}
%and using Lemma~\ref{lem:clustering_error_T_term} we get
%\begin{align*}
%	\Exp{c^\ast(k_n; \Gbox)\ind{A_n^c}\ind{B_n}}
%	&\le \Exp{\frac{(1+\smallO{1})\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%		\ind{A_n^c}} + \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%\end{align*}
%For the last step we use H\"{o}lder's inequality and Proposition \ref{prop:concentration_tilde_T_P_n} to get
%\begin{align*}
%	\Exp{\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\ind{A_n^c}}
%	&\le \Exp{\left(\frac{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}
%		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right)^2}^{\frac{1}{2}} \Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \left(1 + \smallO{1}\right) 	
%		\frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
%		\Prob{A_n^c}^{\frac{1}{2}}\\
%	&= \bigO{\Exp{c^\ast(k_n; \Gbox)}}\Prob{A_n^c}^{\frac{1}{2}} = \smallO{\Exp{c^\ast(k_n; \Gbox)}},
%\end{align*}
%since $\Prob{A_n} = o(1)$. This establishes \eqref{eq:L1_convergence_second_term} and hence we conclude that
%\[
%	\Exp{\left|c^\ast(k_n; \Gbox) - c_{\Pcal, n}(k_n)\right|} = \smallO{\Exp{c_{\Pcal,n}(k_n)^\ast}}.
%\]
%\end{proof}

\begin{proof}[Proof of Proposition \ref{prop:concentration_local_clustering_P_n}]
Again, we write
\begin{align*}
	c^\ast(k_n; \Gbox) = \frac{\widetilde{T}_{\Pcal,n}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	+ \frac{\left(T_{\Pcal,n}(k_n)-\widetilde{T}_{\Pcal,n}(k_n,C)\right)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}},
\end{align*}
so that by~\eqref{eq:asymptotic_equivalence_contraint_triangle_count},
\begin{equation}\label{eq:concentration_local_clustering_P_n_main}
	\Exp{c^\ast(k_n; \Gbox)} = \frac{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	+ \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
\end{equation}
%Therefore, by Lemma \ref{lem:L1_convergence_c_Pcal_n},
%\begin{align*}
%	\Exp{\left|c_{\Pcal,n}(k_n) - \Exp{c^\ast(k_n; \Gbox)}\right|}
%	&\le \Exp{\left|c^\ast(k_n; \Gbox) - \Exp{c^\ast(k_n; \Gbox)}\right|}
%		+ \Exp{\left|c_{\Pcal,n}(k_n) - c^\ast(k_n; \Gbox)\right|}\\
%	&= \frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,\varepsilon) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,\varepsilon)}\right|}}
%		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}} + \smallO{\Exp{c^\ast(k_n; \Gbox)}}.
%		\numberthis \label{eq:concentration_local_clustering_P_n_term1}
%\end{align*}
Next, we use Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,C) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)^2} 
		- \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\Pcal,n}(k_n,C) - \Exp{\widetilde{T}_{\Pcal,n}(k_n,C)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	&= \smallO{\Exp{c^\ast(k_n; \Gbox)}},
\end{align*}
which together with~\eqref{eq:concentration_local_clustering_P_n_main} yields the required result.
\end{proof}

\subsection{Joint neighborhoods and degrees in $\Gbox$}

To prove Proposition~\ref{prop:concentration_tilde_T_P_n} we need to understand the joint degree distribution in $\Gbox$ and subsequently the joint neighborhoods of two points $p, p^\prime \in \Rcal$. We perform the analysis in this section. The two main results, which will be the crucial technical ingredients for the proof of Proposition~\ref{prop:concentration_tilde_T_P_n}, are Lemma~\ref{lem:joint_degree_factorization} and Lemma~\ref{lem:joint_degree_distribution_shift}.

@@

We start with analyzing joint neighborhoods in $\Gbox$. Let $p, p^\prime \in \Rcal$. Then we denote by $\Ncal_{\text{box}}(p\Delta p^\prime)$ the number of disjoint neighbors of $p$ and $p^\prime$ in $\Gbox$, i.e. those points that belong to either $\BallPon{p}$ or $\BallPon{p^\prime}$ but not to their intersection. In addition we denote by $\Ncal_{\text{box}}(p,p^\prime)$ the number of joint neighbors of $p$ and $p^\prime$. We shall establish lower bounds on the expect number of these disjoint neighbors $\Exp{\Ncal_{\text{box}}(p\Delta p^\prime)}$ as well as an asymptotic expression for the expected number of joint neighbors. For this we will distinguish between the cases where the distance between the $x$-coordinates of $p$ and $p^\prime$ is small or large. Figure~\ref{fig:representation_disjoint_neighborhoods_P_n}-\ref{fig:representation_disjoint_neighborhoods_P_n_3} show the different situations that occur.

@@

	%The box \Rcal
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

@@

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
\label{fig:representation_disjoint_neighborhoods_P_n}
\end{figure}~
\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

@@

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $e^{\frac{y + y^\prime}{2}} < |x - x^\prime| \le e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
\label{fig:representation_disjoint_neighborhoods_P_n_2}
\end{figure}~
\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

@@

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x - x^\prime| > e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$ used for the proof of Lemma \ref{lem:disjoint_neighbors_P_n_large}.}
\label{fig:representation_disjoint_neighborhoods_P_n_3}
\end{figure}

We start by analyzing the shape of the neighborhoods. Due to symmetry and the fact that we have identified the left and right boundaries of the box $\Rcal$, we can, without loss of generality, assume that $p = (0,y)$ and $p^\prime = (x^\prime,y^\prime)$ with $x^\prime > 0$ and $y^\prime \le y$. To understand the computation it is helpful to have a picture of the different situations. Figure \ref{fig:representation_disjoint_neighborhoods_P_n} and Figure \ref{fig:representation_disjoint_neighborhoods_P_n_2} show two different situations for small distance in the $x$-coordinates, in which case the number of disjoint neighbors is small. The case where this distance is large, and the number of disjoint neighbors is expected to be large, is show in Figure \ref{fig:representation_disjoint_neighborhoods_P_n_3}. There are several different quantities that are important. The first are the heights $h_1(p^\prime)$ and $h_2(p^\prime)$ where, respectively, the left and right boundaries of the ball $\BallPon{p^\prime}$ go outside the box $\Rcal$. Note that when $x = 0$ then these height are the same and we denote this by $h(y)$. We also need to know the coordinates $h(p,p^\prime)$ and $x^\ast(p,p^\prime)$ of the intersection of the right boundary of the neighborhood of $p$ with the left boundary of the neighborhood of $p^\prime$. Finally we will denote by $d(p,p^\prime)$ the distance between the lower right boundary of $\BallPon{p}$ and the lower left of $\BallPon{p^\prime}$, which is positive only when the bottom parts of both neighborhoods do not intersect, compare Figures \ref{fig:representation_disjoint_neighborhoods_P_n} and \ref{fig:representation_disjoint_neighborhoods_P_n_3}. The full expressions of all these functions are given below for further reference.

@@

\begin{lemma}\label{lem:disjoint_neighbors_P_n}
Let $p, p ^\prime \in \Rcal$. Then whenever $|x - x^\prime| \le e^{\frac{y}{2}} + e^{\frac{y^\prime}{2}}$,
\begin{align*}
	\Exp{\Ncal_{\text{box}}(p \Delta p^\prime)}
	&\ge \frac{\nu}{\pi}|x^\prime - x|\left(1 - \left(\frac{2}{\pi}\right)^{2\alpha}e^{-\alpha(R_n - y^\ast)}\right)\\
	&\hspace{10pt}+ \xi_{\alpha, \nu}\left|e^{\frac{y}{2}} - e^{\frac{y^\prime}{2}}\right|
		\left(1 - \left(\frac{2}{\pi}\right)^{2\alpha - 1}e^{-(\alpha - \frac{1}{2})(R_n-y^\ast)}\right).
\end{align*}
\end{lemma}
